---
title: 'MA Component: Add Bayesian Structure Time Series Model into rapbf'
author: "Justin Ji"
date: "May 15, 2018"
output:
  pdf_document:
    toc: yes
  html_notebook:
    toc: yes
    toc_float: yes
---

# Design

### Overview
Add simple bayesian structure time series model into pick best forecast component.

### Detailed behavior
The mathematical models behind the approach are:

* Observation Equation

$$Time Series = Trend + Seasonality + Regression + Noise$$
$$Y_t = U_t + S_t + \beta^T X_t+ \epsilon_t$$

* State/Transition/Process Equation

$$U_t=U_{t-1}+\delta_{t-1}+\omega_t$$
$$\delta_t=\delta_{t-1}+v_t$$
$$S_t=\sum_{s=1}^{s-1} S_{t-s}+\gamma_t$$
Where $\epsilon_t$, $\omega_t$, $v_t$ and $\gamma_t$ are independent components of Gaussian random noise. $U_t$ is the current level of the trend, the current slop of the trend is $\delta_t$. The seasonal component $S_t$ can be thought of as a set of $s$ of seasons. $X_t$ is the external regressors and $\beta^T$ is the regression coefficients.

If the univariate time series data set is used for forecasting, the Bayesian structural model can be simplified as follow, since the series doesn't have any regressors. 

* Observation Equation

$$Time Series = Trend + Seasonality + Noise$$
$$Y_t = U_t + S_t + \epsilon_t$$

* State/Transition/Process Equation

$$U_t=U_{t-1}+\delta_{t-1}+\omega_t$$
$$\delta_t=\delta_{t-1}+v_t$$
$$S_t=\sum_{s=1}^{s-1} S_{t-s}+\gamma_t$$

### Considerations
* The methodology in the prototype is simple Bayesian time series model for the univariate time series forecasting. 
    - The pros of the prototyped model is that Bayesian time series models are more transparent than ARIMA model. They also facilitate better handling of uncertainty, a key feature when planning for the future. Bayesian time series model fits perfectly with sequential learning and decision making and it directly leads to exact small sample results. 
    - The cons of the prototyped model is that the partial seasonality will be threw into noise/error term. The accurary of the forecasting results might be slighly worse than ARIMA model for some cases.  

* What are not in the prototype ?
    - The model with regressor (i.e. external variable) is not considered. 
    - The input data is partial year data is not considered. 
    - Since there is no project using Bayesian time series model at RA, the model is not covered in any Atlas Toolkit so far. 

# Function Overview

### Function

```{r eval=FALSE}
rabsts(data, season.num, season.dur, pred.horizon)
```
### Input
*             data - Input data.
*             season.num - The number of season to be modeled. 
              i.e., for a time series with quarterly seasonality, season.num=4.
*             season.dur - The number of time periods in each season. 
              i.e., when season.num=4, 
              if it is monthly data, then season.dur=3;                
              if it is weekly data, then season.dur=13.
*             pred.horizon - The number of periods you wish to predict. 
Note: seasonal component (season.num and season.dur) can be determined by external analysis.

### Output

A data frame containing the following as columns:

*             result$mean - The posterior mean of the prediction; 
*             result$median - The posterior median of the prediction.

### Required R Package
*             bsts

# Prototype Code in R

### Function Definition
```{r}
## BEGIN: function ##
rabsts <- function(data, season.num, season.dur, pred.horizon) {
  ##get trend or/and seasonal state
  if(season.num==0){
    ss <- bsts::AddLocalLinearTrend(list(), data)}
  else{
    ss <- bsts::AddLocalLinearTrend(list(), data)
    ss <- bsts::AddSeasonal(ss, data, nseasons = season.num, season.duration = season.dur)}
  
  ##build Bayesial model
  bsts.model <- bsts::bsts(data, state.specification = ss, niter = 666, ping=0, seed=1000)
  
  ##predict
  result<-bsts::predict.bsts(bsts.model, horizon = pred.horizon, burn = SuggestBurn(0.1, bsts.model), quantiles = c(.025, .975))
  pred<-data.frame(result$mean,result$median)
  return(pred)
}
## END: function ##
```

### Example of calling the function with sample data

First, we need to load the required R packages.
```{r}
suppressMessages(library(bsts))
```

Now, we need some input data. The unit test data sets for rapbf package is hired for the prototype. It looks like:
```{r echo=FALSE}
#setwd(choose.dir()) #this is used to set the working directory
inputData<-read.csv("historical.csv") # show input data
holdoutData<-read.csv("holdout.csv") # show holdout data
```

Segment the data by seriesId:
```{r}
#Input data
dat <- subset(inputData, seriesId=="1")
dat <- droplevels(dat)
#Holdout data
holdout.dat<-subset(holdoutData, seriesId=="1")
holdout.dat <- droplevels(holdout.dat)
```

Apply the Bayesial time series model for forecast:

* Seasonal series:
```{r}
x<-rabsts(dat$value,season.num=12,season.dur=1,pred.horizon=3) #assume that there are 12 seasons and each season is one month.
data.frame(x,holdout.dat$value)
```
* Non-seasonal series:
```{r}
y<-rabsts(dat$value,season.num=0,pred.horizon=3) #If no season, set season.num=0.
data.frame(y,holdout.dat$value)
```
